---
title: "Laptop Purchasing Dashboard"
output: 
  html_document:
runtime: shiny
---
```{r include = FALSE}
library(dplyr)
library(ggplot2)
library(gridExtra)
Laptopdf = read.csv("LaptopData.csv")

```
This Laptop Purchasing Dashboard is an interactive tool designed to help users explore and analyze laptop pricing trends and make informed purchasing decisions. By offering a combination of dynamic visualizations and filtering options, the dashboard allows users to examine the distribution of laptop prices, compare different brands, and filter options based on their budget.

Most Common Brand: 
```{r echo = FALSE}
brand_counts <- table(Laptopdf$Company)
most_common_brand <- names(which.max(brand_counts))
print(most_common_brand)
```
Mean Price of Laptops
```{r echo = FALSE}
meanprice = mean(Laptopdf$Price_GBP)
print(meanprice)
```
Total Number of Laptops
```{r echo = FALSE}
TotalLaptops = nrow(Laptopdf)
print(TotalLaptops)
```




```{r echo = FALSE, fig.width=10, fig.height=10}


p1 = ggplot(Laptopdf, aes(x = Price_GBP)) + geom_histogram(binwidth = 100, fill = "blue", color = "black") +labs(title = "Distribution of Laptop Prices", x = "Price (GBP)", y = "Count")
PricePerCompany = Laptopdf %>% group_by(Company) %>% summarise(meanprice = mean(Price_GBP))

p2 = ggplot(data = PricePerCompany ,aes(x = Company, y = meanprice)) + geom_col(fill = "blue", color = "black") + xlab("RAM (GB)")+ ylab("Price (GBP)") + ggtitle("Average Price of Computers Within Each Company") + theme(axis.text.x = element_text(angle = 45, vjust= 0.65))
PricePerRam = Laptopdf %>% group_by(Ram = factor(Ram)) %>% summarise(meanprice = mean(Price_GBP))

p3 = ggplot(data = PricePerRam ,aes(x = Ram, y = meanprice)) + geom_col(fill = "blue", color = "black") + xlab("RAM (GB)")+ylab("Price (GBP)") + ggtitle("Price of Computers Within Each RAM Category")

companycounts = as.data.frame(table(Laptopdf$Company))
p4 = ggplot(companycounts, aes(x = Var1, y = Freq)) + geom_bar(stat = "identity", fill = "blue") + theme(axis.text.x = element_text(angle= 45, hjust = 1)) + xlab("Company") +ylab("NUmber of Laptops") + ggtitle("Number of Laptops `Per` Company")
grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)

```


```{r echo = FALSE}
library(shiny)
ui <- fluidPage(
  titlePanel("Affordable Laptop Finder"),
  numericInput("Budget", "Enter your Budget in GBP - Enter 5500 If No Budget", value = 5500, min = 0, step = 100),
  selectInput("sortOrder", "Sort by Price", choices = c("Ascending" = "asc", "Descending" = "desc")),
  tableOutput("affordablelaptop")
)

server <- function(input, output) {
  output$affordablelaptop <- renderTable({
    affordable <- Laptopdf[Laptopdf$Price_GBP <= input$Budget, ]
    
    if (input$sortOrder == "asc") {
      affordable <- affordable[order(affordable$Price_GBP), ]
    } else {
      affordable <- affordable[order(affordable$Price_GBP, decreasing = TRUE), ]
    }
    
    affordable[, c("Company", "Product", "Price_GBP", "Ram" )]
  })
}

shinyApp(ui = ui, server = server)
```

