---
title: "Laptop Purchasing Dashboard"
output: 
  html_document:
runtime: shiny
---
```{r include = FALSE}
library(dplyr)
library(ggplot2)
library(gridExtra)
library(shiny)
Laptopdf = read.csv("LaptopData.csv")

```
This Laptop Purchasing Dashboard is an interactive tool designed to help users explore and analyze laptop pricing trends and make informed purchasing decisions. By offering a combination of dynamic visualizations and filtering options, the dashboard allows users to examine the distribution of laptop prices, compare different brands, and filter options based on their budget.

Most Common Brand: 
```{r echo = FALSE}
brand_counts <- table(Laptopdf$Company)
most_common_brand <- names(which.max(brand_counts))
print(most_common_brand)
```
Mean Price of Laptops
```{r echo = FALSE}
meanprice = mean(Laptopdf$Price_GBP)
print(meanprice)
```
Total Number of Laptops
```{r echo = FALSE}
TotalLaptops = nrow(Laptopdf)
print(TotalLaptops)
```




```{r echo = FALSE, fig.width=10, fig.height=10}


p1 = ggplot(Laptopdf, aes(x = Price_GBP)) + geom_histogram(binwidth = 100, fill = "blue", color = "black") +labs(title = "Distribution of Laptop Prices", x = "Price (GBP)", y = "Count")
PricePerCompany = Laptopdf %>% group_by(Company) %>% summarise(meanprice = mean(Price_GBP))

p2 = ggplot(data = PricePerCompany ,aes(x = Company, y = meanprice)) + geom_col(fill = "blue", color = "black") + xlab("RAM (GB)")+ ylab("Price (GBP)") + ggtitle("Average Price of Computers Within Each Company") + theme(axis.text.x = element_text(angle = 45, vjust= 0.65))
PricePerRam = Laptopdf %>% group_by(Ram = factor(Ram)) %>% summarise(meanprice = mean(Price_GBP))

p3 = ggplot(data = PricePerRam ,aes(x = Ram, y = meanprice)) + geom_col(fill = "blue", color = "black") + xlab("RAM (GB)")+ylab("Price (GBP)") + ggtitle("Price of Computers Within Each RAM Category")

companycounts = as.data.frame(table(Laptopdf$Company))
p4 = ggplot(companycounts, aes(x = Var1, y = Freq)) + geom_bar(stat = "identity", fill = "blue") + theme(axis.text.x = element_text(angle= 45, hjust = 1)) + xlab("Company") +ylab("NUmber of Laptops") + ggtitle("Number of Laptops `Per` Company")
grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)

```


```{r echo = FALSE}
ui <- fluidPage(
  titlePanel("Affordable Laptop Finder"),
  
  # Use fluidRow to arrange the inputs in a row
  fluidRow(
    # Each column takes up a fraction of the row (e.g., 4 out of 12 columns)
    column(4, selectInput("sortOrder", "Sort by Price", choices = c("Ascending" = "asc", "Descending" = "desc"))),
    column(4, selectInput("searchBy", "Search by", choices = c("Price", "RAM", "Company"))),
    column(4, uiOutput("dynamicInput"))  # Dynamic input based on search criteria
  ),
  
  # Display the filtered table
  tableOutput("affordablelaptop")
)

server <- function(input, output, session) {
  
  # Dynamically change the input field based on the selected search criteria
  output$dynamicInput <- renderUI({
    if (input$searchBy == "Price") {
      numericInput("searchValue", "Enter Maximum Price", value = 5500, min = 0, step = 100)
    } else if (input$searchBy == "RAM") {
      numericInput("searchValue", "Enter Minimum RAM (GB)", value = 8, min = 1, step = 1)
    } else {
      textInput("searchValue", "Enter Company Name", value = "")
    }
  })
  
  # Filter and display the table based on the selected criteria
  output$affordablelaptop <- renderTable({
    
    affordable <- Laptopdf
    
    # Apply filter based on selected criteria
    if (input$searchBy == "Price") {
      affordable <- affordable[affordable$Price_GBP <= input$searchValue, ]
    } else if (input$searchBy == "RAM") {
      affordable <- affordable[affordable$Ram >= as.numeric(input$searchValue), ]
    } else if (input$searchBy == "Company") {
      affordable <- affordable[grepl(input$searchValue, affordable$Company, ignore.case = TRUE), ]
    }
    
    # Sort based on Price
    if (input$sortOrder == "asc") {
      affordable <- affordable[order(affordable$Price_GBP), ]
    } else {
      affordable <- affordable[order(affordable$Price_GBP, decreasing = TRUE), ]
    }
    
    # Display selected columns
    affordable[, c("Company", "Product", "Price_GBP", "Ram" )]
  })
}

shinyApp(ui = ui, server = server)
```

